---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
### Reading the data
```{r}
library(readr)
data.2015 <- read_csv("world-happiness/2015.csv")
head(data.2015)
tail(data.2015)
```

### Filtering the data and making it more comfortable for usage
```{r}
require(dplyr)
library(dplyr)

#As soon as we will have data for different years, I create a fuction for filtering and renaming the columns
filter.data <- function(data.set) {
  data.set <- data.set %>%
    rename(
    happ_score = `Happiness Score`,
    GDP = `Economy (GDP per Capita)`,
    family = Family,
    health = `Health (Life Expectancy)`,
    freedom = Freedom,
    trust = `Trust (Government Corruption)`,
  ) %>%
  select(c(Country, Region, happ_score, GDP, family, health, freedom, trust))
  return(data.set)
}
```

```{r}
#Updated data
#variables left: country, happ_score, GDP, family, health, freedom, trust
data.2015 <-filter.data(data.2015)
```

### Building a correlation matrix to find the most important variables
```{r}
require(corrplot)
library(corrplot)
require(PerformanceAnalytics)
library("PerformanceAnalytics")
data.2015
cor_data <- data.2015[, c(3, 4, 5, 6, 7, 8)]

corr_matrix <- cor(cor_data)
#simple correlation matrix
round(corr_matrix, 2)
#visualized
corrplot(corr_matrix, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

### Correaltion hypothesis testing
```{r}
require(ggplot2)
require(dplyr)
require(plotly)
library(ggplot2)
#Correaltion hypothesis testing
cor.test(data.2015$happ_score, data.2015$GDP)
cor.test(data.2015$happ_score, data.2015$trust)
cor.test(data.2015$happ_score, data.2015$family)
cor.test(data.2015$happ_score, data.2015$health)
cor.test(data.2015$happ_score, data.2015$freedom)

ggplot(data = data.2015, aes(x=data.2015$happ_score)) + geom_histogram(binwidth = 0.3)

t.test(data.2015$happ_score, data.2015$GDP, data = data.2015)
```

The correlation matrix shows us correlation between all the variables, however we are most interested in variables higly correlated with Happiness Score. 
From this approach we can define family, GDP and health system to be significant but let`s use one more approach.
```{r}
data.2015.europe <- data.2015 %>% filter(data.2015[2] == "Western Europe" | data.2015[2] == "Central and Eastern Europe")
data.2015.north.america <- data.2015 %>% filter(data.2015[2] == "North America")
data.2015.australia.newzeland <- data.2015 %>% filter(data.2015[2] == "Australia and New Zealand")
data.2015.middle.east.africa <- data.2015 %>% filter(data.2015[2] == "Middle East and Northern Africa" | data.2015[2] == 'Sub-Saharan Africa')
data.2015.latin.america <- data.2015 %>% filter(data.2015[2] == "Latin America and Caribbean")
data.2015.asia <- data.2015 %>% filter(data.2015[2] == "Southeastern Asia" | data.2015[2] == "Eastern Asia")

#correlation matrix of factors by region
corrplot(cor(data.2015.europe[, c(3, 4, 5, 6, 7, 8)]), order = "hclust",type = "upper", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for Europe")

corrplot(cor(data.2015.north.america[, c(3, 4, 5, 6, 7, 8)]), type = "upper", order = "hclust", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for North America")
corrplot(cor(data.2015.australia.newzeland[, c(3, 4, 5, 6, 7, 8)]), type = "upper", order = "hclust", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for Australia and New Zeland")
corrplot(cor(data.2015.middle.east.africa[, c(3, 4, 5, 6, 7, 8)]), type = "upper", order = "hclust", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for Middle East and Africa")
corrplot(cor(data.2015.latin.america[, c(3, 4, 5, 6, 7, 8)]), type = "upper", order = "hclust", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for Latin America")
corrplot(cor(data.2015.asia[, c(3, 4, 5, 6, 7, 8)]), type = "upper", order = "hclust", mar=c(0,0,1,0), tl.col = "black", tl.srt = 45, title = "Correlation matrix for Asia")
```

### Using Random Forest for variable selection
```{r}
require("mlbench")
require("caret")
library(randomForest)
library('jtools')

model.data <- data.2015[c(3, 4, 5,6, 7, 8)]
#simple linear model
lm.model1 <- lm(happ_score ~., data=model.data)
summary(lm.model1)

#random forest approach
randForest.selection <- randomForest(happ_score ~., data=model.data)
importance(randForest.selection)

summ(lm.model1)
```

From the summary of both algorithms, we should definetely exclude trust feature because it isn`t significant.

### Data visualization 1
```{r}
par(mfrow=c(2,2))
plot(data.2015$GDP, data.2015$happ_score,  xlab= "GDP per capita", ylab = "Happiness Score", main = "2015 GDP vs Happiness Score")
#we see linear relationship between parameters
plot(data.2015$family, data.2015$happ_score, xlab = "Family", ylab= "Happiness Score", main = "2015 Family vs Happiness Score")
#in this case we can spot some outliers, but overall relationship is also linear
plot(data.2015$health, data.2015$happ_score, xlab = "Health", ylab= "Happiness Score", main = "2015 Health vs Happiness Score")

plot(data.2015$freedom, data.2015$happ_score, xlab = "Freedom", ylab= "Happiness Score", main = "2015 Freedom vs Happiness Score")
#not strong correlation

```

### Simple linear models
```{r}
data.2015.lmodelGDP <- lm(happ_score ~ GDP , data = data.2015)
summary(data.2015.lmodelGDP)
y <- data.2015$happ_score
x <- data.2015$GDP
ggplot(data = data.2015, aes(GDP, happ_score)) + geom_point() + geom_smooth(method = lm, formula = y ~ x)

data.2015.lmodelGDP.family <- lm(happ_score ~ GDP + family , data = data.2015)
summary(data.2015.lmodelGDP.family)
y <- data.2015$happ_score
x <- data.2015$GDP + data.2015$family
ggplot(data = data.2015, aes(GDP, happ_score)) + geom_point() + geom_smooth(method = lm, formula = y ~ x)

data.2015.lmodelGDP.family.health <- lm(happ_score ~ GDP + family + health, data = data.2015)
summary(data.2015.lmodelGDP.family.health)
y <- data.2015$happ_score
x <- data.2015$GDP + data.2015$family + data.2015$health
ggplot(data = data.2015, aes(GDP + family + health, happ_score)) + geom_point()  +  geom_smooth(method = lm, formula = y ~ x)

data.2015.lmodel.full <- lm(happ_score ~ GDP + family + health + freedom, data = data.2015)
summary(data.2015.lmodel.full)
y <- data.2015$happ_score
x <- data.2015$GDP + data.2015$family + data.2015$health + data.2015$freedom
ggplot(data = data.2015, aes(GDP + family + health + freedom, happ_score)) + geom_point()  +  geom_smooth(method = lm, formula = y ~ x)

#report-friendly format
export_summs(data.2015.lmodelGDP, data.2015.lmodelGDP.family, data.2015.lmodelGDP.family.health, data.2015.lmodel.full, scale= TRUE,  model.names = c("GDP only", "GDP and family only", "GDP, family and health", "All feat. included"))
```
### Analysis of residuals and test for heteroscedasticity
```{r}
model_fitted <- data.2015$happ_score - residuals(data.2015.lmodel.full)
plot(model_fitted, residuals(data.2015.lmodel.full), ylab="Residuals")
abline(h=0)
```

Here the residuals are distributed more or less equally around the mean and we can't observe a trend that the variance increases as the value increases, so heteroscedasticity is not present here. 
Summary of linear models prove our assumptions about significance of the variables. Adding them one by one improves our model.
Also we observe that when freedom is added to the model, health becomes a little less significant.
However, using this simple linear model I assume that observations are not related with each other within certain groups. 
But, I want to regress the happiness score based on the region, because previously we observed that correlation matrices of features vary from region to region.
So, I perform linear regression on the clustered by region.

This means that we assume independence across clusters but correlation within clusters.

### Clustered Data Linear Model
```{r}
library(lmtest)
library(multiwayvcov)
data.2015.lmodel.full.clustered <- miceadds::lm.cluster(data=data.2015, formula = happ_score ~ GDP + family + health + freedom, cluster = "Region")

coef(data.2015.lmodel.full.clustered)
vcov(data.2015.lmodel.full.clustered)
summary(data.2015.lmodel.full.clustered)

#for comfortable comparisonv
print("Regular model")
summary(data.2015.lmodel.full)
print("Clustered SE")
summary(data.2015.lmodel.full.clustered)
```

After we clustered the data, generally standard error of estimates became lower but std.error of freedom increased, t-statistics is also lower.

### Time-Series Evolution Visualization
```{r}
#load more data
data.2016 <- read_csv("world-happiness/2016.csv")
data.2017 <- read_csv("world-happiness/2017.csv")
data.2018 <- read_csv("world-happiness/2018.csv")
data.2019 <- read_csv("world-happiness/2019.csv")

#data in 2017-2019 does not have Region column so I merge it by country with 2016 dataset and select desired columns
data.2017 <- merge(data.2017, data.2016, by = c("Country"))
data.2017 <- data.2017 %>% rename(
  happ_score = Happiness.Score,
  GDP = Economy..GDP.per.Capita.,
  family = Family.x, 
  health = Health..Life.Expectancy., 
  freedom = Freedom.x
) %>% select(c(Country, Region, happ_score, GDP, family, health, freedom))
data.2017 <-  data.2017[order(-data.2017[3]), ]
#standartize the data
data.2018 <- data.2018 %>% rename(
  Country = `Country or region`)
data.2018 <- merge(data.2018, data.2016, by = c("Country"))
data.2018 <- data.2018 %>% rename(
  happ_score = Score,
  GDP = `GDP per capita`,
  #no family, I add NA
  health = `Healthy life expectancy`,
  freedom = `Freedom to make life choices`
)

data.2019 <- data.2019 %>% rename(
  Country = `Country or region`)
data.2019 <- merge(data.2019, data.2016, by = c("Country"))
data.2019 <- data.2019 %>% rename(
  happ_score = Score,
  GDP = `GDP per capita`,
  #no family, I add NA in a loop
  health = `Healthy life expectancy`,
  freedom = `Freedom to make life choices`
)
data.2016 <- filter.data(data.2016)
data.lst <- list(data.2016, data.2017, data.2018, data.2019)
top.lst <- list();
full.lst <- list();
#filter, clean up and select top countries 
i <- 1
for (data.set_i in 1:length(data.lst)) {
  if ("family" %in% colnames(data.lst[[data.set_i]])) {
    new.set <- data.lst[[data.set_i]] %>% select(c(Country, Region, happ_score, GDP, family, health, freedom))
  } else {
    new.set <- data.lst[[data.set_i]] %>% select(c(Country, Region, happ_score, GDP, health, freedom))
    new.set$family = NA
    
  }
  new.set$year = toString(2015 + i)
  top.lst[[i]] <- new.set %>% filter(new.set[3] > 7)
  full.lst[[i]] <- new.set
  i <- i + 1
}

top.2015 <- data.2015 %>% filter(data.2015[3] > 7) %>% select(c(Country, Region, happ_score, GDP, family, health, freedom))
top.2015$year = "2015"

#merged set over years with top counties whose happiness score was over 7
merged.set.top.countries <- rbind(top.2015, top.lst[[1]], top.lst[[2]], top.lst[[3]], top.lst[[4]])
head(merged.set.top.countries)
```

```{r}
require(ggplot2)
require(dplyr)
require(plotly)
library(ggplot2)
library(dplyr)
library(plotly)
p <- merged.set.top.countries %>% ggplot(aes(x = year, y = happ_score, colour = Country, group = Country)) + geom_line() + ylab("Happiness Score") + xlab("Year")
p

```

We see how the happiness rate changed over years and I decided to examine the most interesting cases separately.
We see a big decline in happiness score in Austria in 2017 which was then followed by dramatic increase.
Also, the United States has completely missing data on the graph in the time range between 2016 and 2018 which means that its value dropped under 7.0.

## Dramatic Cases Examination
### Austria
```{r}
austria.time.info <- merged.set.top.countries %>% filter(merged.set.top.countries[1] == "Austria")
austria.time.info

par(mfrow=c(2,2))
plot(austria.time.info$year, austria.time.info$happ_score, type="o", xlab= "Year", ylab = "Happiness Score")
plot(austria.time.info$year, austria.time.info$GDP, type="o", xlab= "Year", ylab = "GDP")
plot(austria.time.info$year, austria.time.info$health, type="o", xlab= "Year", ylab = "Health")
plot(austria.time.info$year, austria.time.info$freedom, type="o", xlab= "Year", ylab = "Freedom")
```

### United States
```{r}
data.2015$year = "2015"

#merge set with the rest of counties out of top list
merged.set.2 <- rbind(data.2015 %>% select(-c(trust)), full.lst[[1]], full.lst[[2]], full.lst[[3]], full.lst[[4]])
united.states.info <- merged.set.2 %>% filter(merged.set.2[1] == "United States")

par(mfrow=c(2,2))
plot(united.states.info$year, united.states.info$happ_score, type="o", xlab= "Year", ylab = "Happiness Score")
plot(united.states.info$year, united.states.info$GDP, type="o", xlab= "Year", ylab = "GDP")
plot(united.states.info$year, united.states.info$health, type="o", xlab= "Year", ylab = "Health")
plot(united.states.info$year, united.states.info$freedom, type="o", xlab= "Year", ylab = "Freedom")

```

## Finland
```{r}
finland.info <- merged.set.2 %>% filter(merged.set.2[1] == "Finland")

par(mfrow=c(2,2))
plot(finland.info$year, finland.info$happ_score, type="o", xlab= "Year", ylab = "Happiness Score")
plot(finland.info$year, finland.info$GDP, type="o", xlab= "Year", ylab = "GDP")
plot(finland.info$year, finland.info$health, type="o", xlab= "Year", ylab = "Health")
plot(finland.info$year, finland.info$freedom, type="o", xlab= "Year", ylab = "Freedom")

```

From now on, we can state that GDP and happiness score are not perfectly correlated and it does not always explain the happiness rate in the country. While health almost always has the same curve as happiness rate.

So, now I suggest adding to our model some factors which describe cultural life of the country and see if they improve the model.

```{r}
library(readr)
library(dplyr)
#read primary education data
primary.educ <- read.csv("enrolment_primary_educ.csv")
primary.educ <- primary.educ %>% filter(primary.educ[6] > 2014 & primary.educ[6] <=2017)
#read population data and format it
population <- read_csv("population.csv")
population <- population %>% select(c(`Country Name`,`2015`, `2016`, `2017`))
population <- population[rep(seq_len(nrow(population)), each=3), ]
#add years as rows
population$year = c(2015, 2016, 2017)

population.copy <- population[c(2, 3, 4)]

#fucntion for chossing every y row
nrowgen <- function(x, y) {
  n <- nrow(x)
  b <- seq(1, n, y)
  r <- length(b)
  c <- data.frame()
  for ( i in 1:r) {
    abc <- x[b[i], ]
    c <- rbind(c, abc)
  }
  return(c)
}
population.values <- nrowgen(population.copy, 3)

#adding all population numbers to one vector
population.values.vector <- c();
names(population.values) <- NULL
for (i in 1:nrow(population.values)) {
  population.values.vector <- c(population.values.vector,unlist(c(population.values[i, ])) )
}
population$pop_number = population.values.vector
population <- population %>% select(c(`Country Name`, year, pop_number))
primary.educ <- primary.educ %>% select(c(Indicator, Country, Time, Value))

population <- population %>% rename(
  Country = `Country Name`,
  Time = year
)
#final population dataset, formatted
#merged population with primary educ
primary.educ.merged <- merge(primary.educ, population, by=c("Country", "Time"))
```
```{r}
secondary.educ <- read_csv("enrolment_secondary_educ.csv")
secondary.educ <- secondary.educ %>% filter(secondary.educ[6] > 2014 & secondary.educ[6] <=2017)
secondary.educ <- secondary.educ %>% select(c(Indicator, Country, Time, Value))
secondary.educ.merged <- merge(secondary.educ, population, by=c("Country", "Time"))
```
```{r}
tertiary.educ <- read_csv("enrolment_tertiary_educ.csv")
tertiary.educ <- tertiary.educ %>% filter(tertiary.educ[6] > 2014 & tertiary.educ[6] <=2017)
tertiary.educ <- tertiary.educ %>% select(c(Indicator, Country, Time, Value))
tertiary.educ.merged <- merge(tertiary.educ, population, by=c("Country", "Time"))
```
```{r}
#enrolment in primary education for 2015
primary.educ.merged.2015 <- primary.educ.merged %>% filter(primary.educ.merged[2] == 2015)
data.2015 <- merge(data.2015, primary.educ.merged.2015[c(1, 4, 5)], by=c("Country"))
data.2015 <- data.2015 %>% rename(
  value_primary = Value
)
data.2015$ratio_primary <- data.2015[, "value_primary"] / data.2015[, "pop_number"]

secondary.educ.merged.2015 <- secondary.educ.merged %>% filter(secondary.educ.merged[2] == 2015)
data.2015 <- merge(data.2015, secondary.educ.merged.2015[c(1, 4, 5)], by=c("Country"))
data.2015 <- data.2015 %>% rename(
  value_secondary = Value
) %>% select(-c(pop_number.y))
data.2015$ratio_secondary <- data.2015[, "value_secondary"] / data.2015[, "pop_number.x"]

tertiary.educ.merged.2015 <- tertiary.educ.merged %>% filter(tertiary.educ.merged[2] == 2015)
data.2015 <- merge(data.2015, tertiary.educ.merged.2015[c(1, 4, 5)], by=c("Country"))
data.2015 <- data.2015 %>% rename(
  value_tertiary = Value
) %>% select(-c(pop_number))
data.2015$ratio_tertiary <- data.2015[, "value_tertiary"] / data.2015[, "pop_number.x"]

```

### Linear Model including education feature
```{r}
data.2015.lmodel.full.educ <- lm(happ_score ~ GDP + family + health + freedom + ratio_primary + ratio_secondary + ratio_tertiary , data = data.2015)
summary(data.2015.lmodel.full.educ)
y <- data.2015$happ_score
x <- data.2015$GDP + data.2015$family + data.2015$health + data.2015$freedom + data.2015$ratio_primary + data.2015$ratio_secondary + data.2015$ratio_tertiary
ggplot(data = data.2015, aes(GDP + family + health + freedom + ratio_primary + ratio_secondary + ratio_tertiary, happ_score)) + geom_point()  +  geom_smooth(method = lm, formula = y ~ x)


```

### Clustered Model including enrolment in education
```{r}
library(lmtest)
library(multiwayvcov)
data.2015.lmodel.full.clustered.educ <- miceadds::lm.cluster(data=data.2015, formula = happ_score ~ GDP + family + health + freedom + ratio_primary + ratio_secondary , cluster = "Region")

coef(data.2015.lmodel.full.clustered.educ)
vcov(data.2015.lmodel.full.clustered.educ)

summary(data.2015.lmodel.full.clustered.educ)



model_fitted <- data.2015$happ_score - residuals(data.2015.lmodel.full.educ)
plot(model_fitted, residuals(data.2015.lmodel.full.educ))
abline(h=0)

```

## Pooled Data
```{r}
# fuction for adding education to the dataset
educ.to.dataset <- function(data.set, year) {
  primary.educ.merged.new <- primary.educ.merged %>% filter(primary.educ.merged[2] == year)
  primary.educ.merged.new
  data.set <- merge(data.set, primary.educ.merged.new[c(1, 4, 5)], by=c("Country"))
  data.set <- data.set %>% rename(
    value_primary = Value
  )
  data.set$ratio_primary <- data.set[, "value_primary"] / data.set[, "pop_number"]
  
  secondary.educ.merged.new <- secondary.educ.merged %>% filter(secondary.educ.merged[2] == year)
  secondary.educ.merged.new
  data.set <- merge(data.set, secondary.educ.merged.new[c(1, 4, 5)], by=c("Country"))
  data.set <- data.set %>% rename(
    value_secondary = Value
  ) %>% select(-c(pop_number.y))
  data.set$ratio_secondary <- data.set[, "value_secondary"] / data.set[, "pop_number.x"]
  
  tertiary.educ.merged.new <- tertiary.educ.merged %>% filter(tertiary.educ.merged[2] == year)
  tertiary.educ.merged.new
  data.set <- merge(data.set, tertiary.educ.merged.new[c(1, 4, 5)], by=c("Country"))
  data.set
  data.set <- data.set %>% rename(
    value_tertiary = Value
  ) %>% select(-c(pop_number))
  data.set$ratio_tertiary <- data.set[, "value_tertiary"] / data.set[, "pop_number.x"]
  data.set
  return(data.set)
}
```

```{r}
data.2016 <- educ.to.dataset(data.2016, 2016)
```

```{r}
data.2017$year = 2017
data.2017 <- educ.to.dataset(data.2017, 2017)
```

### Science Development feature
```{r}
science_development <- read_csv("culture/science_development.csv")
science_development <- science_development %>% filter(science_development[5] == 2015 | science_development[5] == 2016 | science_development[5] == 2017)
science_development <- science_development %>% filter(science_development[2] == 'Total R&D personnel (FTE) - Total')
                                                      
science_development <- science_development %>% rename(
  year = Time,
  value_science_develop = Value) %>% select(c(Country, year, value_science_develop))

science.development.2015 <- science_development %>% filter(science_development[2] == 2015)
data.2015 <- merge(data.2015, science.development.2015[c(1, 3)], by="Country")
data.2015$ratio_science_develop <- data.2015[, "value_science_develop"] / data.2015[, "pop_number.x"]

science.development.2016 <- science_development %>% filter(science_development[2] == 2016)
data.2016 <- merge(data.2016, science.development.2016[c(1, 3)], by="Country")
data.2016$ratio_science_develop <- data.2016[, "value_science_develop"] / data.2016[, "pop_number.x"]

science.development.2017 <- science_development %>% filter(science_development[2] == 2017)
data.2017 <- merge(data.2017, science.development.2017[c(1, 3)], by="Country")
data.2017$ratio_science_develop <- data.2017[, "value_science_develop"] / data.2017[, "pop_number.x"]

```
```{r}

pooled.happ_score.data <- merge(data.2015, data.2016, by=c("Country"))
pooled.happ_score.data <- pooled.happ_score.data %>% rename(
  happ_score.2015 = happ_score.x,
  happ_score.2016 = happ_score.y,
  GDP.2015 = GDP.x,
  GDP.2016 = GDP.y,
  family.2015 = family.x,
  family.2016 = family.y,
  health.2015 = health.x,
  health.2016  = health.y,
  freedom.2015 = freedom.x,
  freedom.2016 = freedom.y,
  trust.2015 = trust.x,
  trust.2016 = trust.y,
  ratio_primary.2015 = ratio_primary.x,
  ratio_primary.2016 = ratio_primary.y,
  ratio_secondary.2015 = ratio_secondary.x,
  ratio_secondary.2016 = ratio_secondary.y,
  ratio_tertiary.2015 = ratio_tertiary.x,
  ratio_tertiary.2016 = ratio_tertiary.y,
  ratio_science_develop.2015 = ratio_science_develop.x,
  ratio_science_develop.2016 = ratio_science_develop.y
) %>% select(-c(Region.y, value_primary.x, value_secondary.x, value_tertiary.x, 
                value_primary.y, value_secondary.y, value_tertiary.y, pop_number.x.y, value_science_develop.x, value_science_develop.y))


pooled.happ_score.data <- merge(pooled.happ_score.data, data.2017 , by=c("Country"))
pooled.happ_score.data <- pooled.happ_score.data %>% rename(
  happ_score.2017 = happ_score,
  GDP.2017 = GDP,
  family.2017 = family,
  health.2017 = health,
  freedom.2017 = freedom,
  ratio_primary.2017 = ratio_primary,
  ratio_secondary.2017 = ratio_secondary,
  ratio_tertiary.2017 = ratio_tertiary,
  ratio_science_develop.2017 = ratio_science_develop
) %>% select(-c(Region, value_primary, value_secondary, value_tertiary, pop_number.x, trust.2015, trust.2016, value_science_develop))
pooled.happ_score.data <- pooled.happ_score.data[rep(seq_len(nrow(pooled.happ_score.data)), each=3), ]
pooled.happ_score.data$year <- c(2015, 2016, 2017)
```

```{r}
happ_scores.vector <- c();
i <- 1
while (i <= nrow(pooled.happ_score.data)) {
  happ_scores.vector <- c(happ_scores.vector,  pooled.happ_score.data[i, 3], pooled.happ_score.data[i, 14], pooled.happ_score.data[i, 23])
  i <- i + 3
}

pooled.happ_score.data$happ_score_pooled = happ_scores.vector

i <- 1
gdp.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  gdp.vector <- c(gdp.vector,  pooled.happ_score.data[i, 4], pooled.happ_score.data[i, 15], pooled.happ_score.data[i, 24])
  i <- i + 3
}
pooled.happ_score.data$gdp_pooled = gdp.vector

i <- 1
family.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  family.vector <- c(family.vector,  pooled.happ_score.data[i, 5], pooled.happ_score.data[i, 16], pooled.happ_score.data[i, 25])
  i <- i + 3
}

pooled.happ_score.data$family_pooled = family.vector

i <- 1
health.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  health.vector <- c(health.vector,  pooled.happ_score.data[i, 6], pooled.happ_score.data[i, 17], pooled.happ_score.data[i, 26])
  i <- i + 3
}

pooled.happ_score.data$health_pooled = health.vector

i <- 1
freedom.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  freedom.vector <- c(freedom.vector,  pooled.happ_score.data[i, 7], pooled.happ_score.data[i, 18], pooled.happ_score.data[i, 27])
  i <- i + 3
}
pooled.happ_score.data$freedom_pooled = freedom.vector

i <- 1
ratio_primary.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  ratio_primary.vector <- c(ratio_primary.vector,  pooled.happ_score.data[i, 10], pooled.happ_score.data[i, 19], pooled.happ_score.data[i, 29])
  i <- i + 3
}
pooled.happ_score.data$ratio_primary_pooled = ratio_primary.vector

i <- 1
ratio_secondary.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  ratio_secondary.vector <- c(ratio_secondary.vector,  pooled.happ_score.data[i, 11], pooled.happ_score.data[i, 20], pooled.happ_score.data[i, 30])
  i <- i + 3
}
pooled.happ_score.data$ratio_secondary_pooled = ratio_secondary.vector

i <- 1
ratio_tertiary.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  ratio_tertiary.vector <- c(ratio_tertiary.vector,  pooled.happ_score.data[i, 12], pooled.happ_score.data[i, 21], pooled.happ_score.data[i, 31])
  i <- i + 3
}

pooled.happ_score.data$ratio_tertiary_pooled = ratio_tertiary.vector


i <- 1
ratio_science_develop.vector <- c();
while (i <= nrow(pooled.happ_score.data)) {
  ratio_science_develop.vector <- c(ratio_science_develop.vector,  pooled.happ_score.data[i, 13], pooled.happ_score.data[i, 22], pooled.happ_score.data[i, 32])
  i <- i + 3
}

pooled.happ_score.data$ratio_science_develop_pooled = ratio_science_develop.vector


pooled.happ_score.data <- pooled.happ_score.data %>% select(c(Country, Region.x, year, happ_score_pooled, gdp_pooled, family_pooled, health_pooled, freedom_pooled, ratio_primary_pooled,ratio_secondary_pooled, ratio_tertiary_pooled, ratio_science_develop_pooled))

```

### Explore through vizualization
```{r}
plot(pooled.happ_score.data$gdp_pooled, pooled.happ_score.data$happ_score_pooled,  xlab= "GDP per capita", ylab = "Happiness Score", main = "Pooled GDP vs Happiness Score")
#we see linear relationship between parameters
plot(pooled.happ_score.data$family_pooled, pooled.happ_score.data$happ_score_pooled,  xlab= "Family", ylab = "Happiness Score", main = "Pooled Family vs Happiness Score")
plot(pooled.happ_score.data$health_pooled, pooled.happ_score.data$happ_score_pooled,  xlab= "Health", ylab = "Happiness Score", main = "Pooled Health vs Happiness Score")
plot(pooled.happ_score.data$ratio_science_develop_pooled, pooled.happ_score.data$happ_score_pooled,  xlab= "Science Development", ylab = "Happiness Score", main = "Pooled Science Develop vs Happiness Score")
```


## Panel models

### Model with standart features given in original dataset
```{r}
library(plm)
library(lmtest)
library(multiwayvcov)

pooled.ols.standardfeat <- plm(formula= pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) +
                    pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled
                  , data=pooled.happ_score.data, model="pooling", index=c("Country", "year"))
summary(pooled.ols.standardfeat)
```


### Full model with education and science development features
```{r}
library(plm)
library(lmtest)
library(multiwayvcov)
pooled.ols.full <- plm(formula= pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) +
                    pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled
                  , data=pooled.happ_score.data, model="pooling", index=c("Country", "year"))
summary(pooled.ols.full)

#adding dummy variables
pooled.happ_score.data$year2016_d <- ifelse(pooled.happ_score.data$year == 2016, 1, 0)
pooled.happ_score.data$year2017_d <- ifelse(pooled.happ_score.data$year == 2017, 1, 0)

#pooled with dummies
pooled.ols.dummies <- plm(formula= pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) + pooled.happ_score.data$year2016_d + pooled.happ_score.data$year2017_d + 
                    pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled
                  , data=pooled.happ_score.data, model="pooling", index=c("Country", "year"))
summary(pooled.ols.dummies)
#dummies are insignificant

```

### Full model. FE and RE and Haussman Test
```{r}
library(plm)
library(lmtest)
library(multiwayvcov)
library(wooldridge)

#clustered by group
coeftest(pooled.ols.full, vcov=vcovHC(pooled.ols.full, type="sss", cluster="group"))

#FE model
fe.model <- plm(formula=pooled.happ_score.data$happ_score_pooled ~
                    pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="within", index=c("Country", "year"))
summary(fe.model)

#RE model
re.model <- plm(formula=pooled.happ_score.data$happ_score_pooled  ~factor(pooled.happ_score.data$Region.x) + 
                    pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="random", index=c("Country", "year"))
summary(re.model)

#Hausmann Test: null hypothesis: preffered RE
phtest(fe.model, re.model)
# in this case we fail to reject null hypothesis
```

### Adding income as a regressor
```{r}
income.data <- read_csv("DP_LIVE_23042020144413075.csv")
income.data $Country = NA
income.data$Country[income.data$LOCATION == "AUS"] <-"Australia"
income.data$Country[income.data$LOCATION == "BEL"] <-"Belgium"
income.data$Country[income.data$LOCATION == "AUT"] <-"Austria"
income.data$Country[income.data$LOCATION == "AUT"] <-"Austria"
country.codes <- c("AUS", "BEL", "AUT", "CAN", "CZE", "DNK", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL",
                   "ITA", "JPN", "LUX", "MEX", "NLD", "NOR", "POL", "PRT", "SVK", "ESP",
                   "SWE", "CHE", "EST", "SVN", "ISL", "CHL", "LVA", "LTU")

country.names <- c("Australia", "Belgium", "Austria", "Canada", "Czechia", "Denmark", "Finland", 
                   "France", "Germany", "Greece", "Hungary", "Ireland", "Italy", "Japan", "Luxembourgh",
                   "Mexico", "Netherlands", "Norway", "Poland", "Portugal", "Slovakia", "Spain", "Sweden", "Switzerland", "Estonia", "Slovenia", "Iceland", "Chile", "Latvia", "Lithuania")

for (i in 1:length(country.codes)) {
  income.data$Country[income.data$LOCATION == country.codes[[i]]] <- country.names[[i]]
}
income.data <- income.data %>% select(c(Country, TIME, Value)) %>% rename(year = TIME, annual_income = Value)
```
```{r}
pooled.happ_score.data <- merge(pooled.happ_score.data, income.data, by=c("Country", "year"))
```

### Pooled OLS with all features and income instead of GDP
```{r}
pooled.ols <- plm(formula= pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) +
                    pooled.happ_score.data$annual_income + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled
                  , data=pooled.happ_score.data, model="pooling", index=c("Country", "year"))
summary(pooled.ols)

#by group
coeftest(pooled.ols, vcov=vcovHC(pooled.ols, type="sss", cluster="group"))
```

### FE model with all features and income instead of GDP
```{r}
#FE model

fe.model <- plm(formula=pooled.happ_score.data$happ_score_pooled ~
                    pooled.happ_score.data$annual_income + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="within", index=c("Country", "year"))
summary(fe.model)
```

### RE model with all features and income instead of GDP
```{r}
#RE model
re.model <- plm(formula=pooled.happ_score.data$happ_score_pooled  ~factor(pooled.happ_score.data$Region.x) + 
                    pooled.happ_score.data$annual_income + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="random", index=c("Country", "year"))
summary(re.model)
```
### Hausmann Test
```{r}
#Hausmann Test
phtest(fe.model, re.model)
```
### Decide between RE regression and simple OLS
```{r}
plmtest(pooled.ols, type=c("bp"))
#use RE
```

### Pooled OLS with all features and income and GDP
```{r}
pooled.ols <- plm(formula= pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) +
                    pooled.happ_score.data$annual_income + + pooled.happ_score.data$gdp_pooled +
                  pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled
                  , data=pooled.happ_score.data, model="pooling", index=c("Country", "year"))
summary(pooled.ols)

#by group
coeftest(pooled.ols, vcov=vcovHC(pooled.ols, type="sss", cluster="group"))
```

### FE model with all features and income and GDP
```{r}
#FE model
fe.model <- plm(formula=pooled.happ_score.data$happ_score_pooled ~ factor(pooled.happ_score.data$Region.x) +
                    pooled.happ_score.data$annual_income + pooled.happ_score.data$gdp_pooled + pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="within", index=c("Country", "year"))
summary(fe.model)
```

### RE model with all features and income and GDP
```{r}
#RE model
re.model <- plm(formula=pooled.happ_score.data$happ_score_pooled  ~factor(pooled.happ_score.data$Region.x) + 
                    pooled.happ_score.data$annual_income + pooled.happ_score.data$gdp_pooled +  pooled.happ_score.data$family_pooled + pooled.happ_score.data$health_pooled + pooled.happ_score.data$freedom_pooled + pooled.happ_score.data$ratio_primary_pooled + pooled.happ_score.data$ratio_secondary_pooled + pooled.happ_score.data$ratio_tertiary_pooled + pooled.happ_score.data$ratio_science_develop_pooled , data=pooled.happ_score.data, model="random", index=c("Country", "year"))
summary(re.model)
```

